---
title: "Machine Learning Project"
author: "Hasnat Ali"
date: "10 sept 2020"
output: html_document
---

The main goal of the Project is to quantify how well an individual perform for a particular activity. This will be accomplished by training a prediction model on the accelerometer data. The algorithm that we will be using for this exercise will be a random forest classifier.

We will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. People were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

The training data and the test data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

####Load libraries

```{r}
library(doMC)
registerDoMC(cores = 7)
library(randomForest)
library(caret)
```


####Download and load data
```{r, echo=FALSE}
train_url<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(train_url, destfile="training.csv", method="curl")
download.file(test_url, destfile="testing.csv", method="curl")
training<-read.csv("training.csv", na.strings=c("NA","#DIV/0!",""))
predictDataset<-read.csv("testing.csv",na.strings=c("NA","#DIV/0!",""))
```

####Clean data

First we remove the columns that aren't the predictor variables
```{r, cache=TRUE}
col.rm <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")
training.rm <- which(colnames(training) %in% col.rm)
training <- training[, -training.rm]
```

Next, we write a custom function that removes all collumns with NA values
```{r, cache=TRUE}
for (colName in names(training)){
  if(class(training[[colName]]) %in% c("integer", "numeric")){
    m<-mean(training[[colName]], na.rm = TRUE)
    training[[colName]][is.na(training[[colName]])]<-m
  }
}
```

In last cleaning step we remove the columns where variance is near zero.
```{r, cache=TRUE}
training<-training[,-nearZeroVar(training)]
```

####Data partitioning
Split the data into training and testing datasets.
```{r, cache=TRUE}
index <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
trainingData <- training[ index, ]
testingData <-  training[ -index, ]


train_control <- trainControl(method="cv", number=3)
model <- train(classe~., data=trainingData, method="rf", trControl = train_control)
```

####Training
We'll using Random Forest algorithm with 3 times cross validation and then choose a best model.
```{r, cache=TRUE}
train_control <- trainControl(method="cv", number=3)
model <- train(classe~., data=trainingData, method="rf", trControl = train_control)
model
```

####Prediction
First we'll make predictions on training dataset and then on testing dataset.
```{r, cache=TRUE}
prediction <- predict(model, trainingData, type="raw")
c<-confusionMatrix(prediction, trainingData$classe)
print("In Sample Error Rate")

prediction <- predict(model, testingData, type="raw")
c<-confusionMatrix(prediction, testingData$classe)

prediction <- predict(model, predictDataset, type="raw")

print(1-c[["overall"]][["Accuracy"]])

prediction <- predict(model, predictDataset, type="raw")
```

####Generating files to coursera submitt
```{r, cache=TRUE}
for (i in seq(20)){
  fileName<-paste("problem",i,".txt",sep="_")
  write.table(prediction[i],file=fileName,quote=FALSE,row.names=FALSE,col.names=FALSE)
}
```
